Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association

### 2022-04-30 papers/ms.tex
update ms

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new window
  on open cluster dispersal.  Here we present an analysis of a group
  of stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
  hereafter the Cep-Her complex.  The group includes four Kepler
  Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in part of the Cep-Her complex
  called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.  Here
  we focus on the former three systems, which are in other regions of
  the association.  Based on kinematic evidence from Gaia, stellar
  rotation periods from TESS, and spectroscopy, these three objects
  are also $\approx$40 million years (Myr) old.  More specifically, we
  find that Kepler-1643 is $46^{+9}_{-7}$\,Myr old, based on its
  membership in a dense sub-cluster of the complex called RSG-5.
  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
  diffuse region that we call CH-2.  Based on the transit shapes and
  high resolution imaging, all three objects are most likely planets,
  with false positive probabilities of $6\times10^{-9}$,
  $4\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
  and KOI-7913 respectively.  These planets demonstrate
  that mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages
  of 40 million years.
\end{abstract}

@@

%
% KOI match numbers from calc_CepHer_group_numbers.py
%
Our fourth and final step was to cross-match the candidate Cep-Her
member list against all known Kepler Objects of Interest.  We used the
Cumulative KOI table from the NASA Exoplanet Archive from 27 March
2022, and also compared against the \texttt{q1\_q17\_dr25} table
\citep{thompson_planetary_2018}.  From the candidate members with
weights exceeding 0.02, this yielded 11 known false positives, 6
``confirmed'' planets, and 8 ``candidate'' planets.  Inspection of the
Kepler data validation summaries and Robovetter classifications for
these objects showed whether they were potentially consistent with
being {\it i)} planets, and {\it ii)} $\lesssim 10^8$ years old, based
on the presence of rotational modulation at the expected period and
amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}.  Four
objects remained after this inspection: Kepler-1627, Kepler-1643,
KOI-7368, and KOI-7913.

@@

\subsection{KOI-7368}
\paragraph{Spectra}
For KOI-7368, we acquired a spectrum on 2015 June 1 using the echelle
spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
Stellar Parameter Classification pipeline for TRES has been described
by \citet{2021tsc2.confE.124B}.  It is based on the synthetic template
library constructed by \citet{Buchhave2012}.  The resulting stellar
parameters ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
cluster-isochrone method within $1$-$\sigma$, though the effective
temperature was more precise by a factor of three.  Auxiliary
spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
\pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm 0.50\,\mkms$,
and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
The Li 6708\AA\ EW measurement procedure yielded ${\rm EW}_{\rm Li} =
236^{+16}_{-14}$\,m\AA.

@@

The theoretical expectation is that mini-Neptunes with sizes of 2 to
3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
This expectation is tied to inferences about the initial distributions
of planetary core mass, core composition, and atmospheric mass
fraction \citep{owen_evaporation_2017}.  The Kelvin-Helmholtz cooling
timescale, which is tied to the entropy of the planetary interior
shortly after disk dispersal, also plays a significant role
\citep{owen_constraining_2020}.  As an example,
\citet{rogers_unveiling_2021} predicted that given a core-mass
distribution peaked at $\approx$$4$\,$M_\oplus$, an ice-poor rock/iron
core composition, and a typical H/He mass fraction of $\approx$4\%,
there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
at times between 10 and 100\,Myr.  The models advanced by
\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
their differences lie in the mechanism for producing the radius
valley.

Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837, and TOI-1227
have sizes that are anomalously large relative to the predicted peak
in planet occurrence at 2 to 3\,$R_\oplus$.  However, their large
sizes can be accommodated by invoking any of {\it i)} larger core
masses, {\it ii)} more volatile-rich compositions, {\it iii)}  larger
initial atmospheric mass fractions, or {\it iv)} longer thermal
cooling times.  Secure mass measurements would help constrain this
parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
semi-amplitudes makes measuring the Doppler orbits very difficult.
\citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
demonstrate that at least some planets at 40\,Myr have sizes that are
consistent with theoretical expectations for mini-Neptunes.  While
selection effects imposed by spot-induced photometric variability are
a likely explanation for why planets this small have not previously
been identified \citep[{e.g.},][]{zhou_2021_tois}, future work should
quantify this bias more carefully, in order to enable empirical
studies of how the planetary size distribution changes at early times.

@@

We have shown that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
to 50 million years old, and that each system is most likely
planetary.  The evidence for the planetary interpretation comes from
an application of \texttt{VESPA} to the Kepler data, alongside new
imaging from NIRC2.  The validity of the \texttt{VESPA} framework
rests on the premise that non-astrophysical false positives can be
rejected.  Kepler-1643 and KOI-7368 both pass all but one of the
canonical vetting tests: the check for centroid offsets between the in
and out-of-transit phases.  For both cases, the weak observed shifts
are consistent with being caused by starspot-induced variability in
specific quarters spuriously moving the stellar center-of-light.
Independently, our imaging rules out companion stars with the
brightness and positions that would be needed to otherwise explain the
reported shifts.

@@

%\clearpage
\acknowledgements
\raggedbottom
% Cycle 3 and 4
L.G.B{.} is supported by the Heising-Simons Foundation 51
Pegasi~b Fellowship and the TESS GI Program (NASA grants 80NSSC21K0335
and 80NSSC22K0298).
R.K{.} is supported by the Heising-Simons Foundation.
J.L.C{.} is supported by NSF AST-2009840 and the TESS GI Program (NASA
grant 80NSSC22K0299).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%%
%We thank the TESS architects
%%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
%and the many TESS team members for their
%efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
%Finally, this research has made use of the Keck Observatory Archive (KOA),
%which is operated by the W. M. Keck Observatory and the NASA Exoplanet
%Science Institute (NExScI), under contract with the National
%Aeronautics and Space Administration.
%We also thank the Keck Observatory staff for their support of
%HIRES and remote observing.  We recognize the importance that the
%summit of Maunakea has within the indigenous Hawaiian community, and
%are deeply grateful to have the opportunity to conduct observations
%from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\paragraph{Table~2} contains candidate Cep-Her members with weights
$D>0.02$ observed by Kepler.  The final catalog of candidate
Cep-Her members will be provided by R.~Kerr et al.\ in prep; Table~2 is
from a preliminary version of that work.  Note that more
restrictive weight cuts should be imposed should one wish to remove
the majority of field star interlopers; at least half of the stars
included in the table are unlikely to be bonafide Cep-Her members.
Table~2  was created by cross-matching candidate Cep-Her members
(selected using Gaia EDR3; Section~\ref{subsec:members}) against a
Kepler to Gaia DR2 cross-match (the \texttt{gaia-kepler.fun}
crossmatch database created by Megan Bedell).  The
\texttt{kic\_dr2\_ang\_dist} column is from the latter table.  The
EDR3 to DR2 match was performed using the
\texttt{gaiaedr3.dr2\_neighbourhood} table, and the closest proper
motion and epoch-corrected angular distance neighbor was taken as the
single best match.  The \texttt{edr3\_dr2\_mag\_diff} column gives
some indication of the reliability of this EDR3 to DR2 conversion, as
there are a few cases between Gaia DR2 and EDR3 where partially
resolved binaries became fully resolved.

\paragraph{Candidate matches between Cep-Her and the Kepler Objects of
Interest:}
The full list of candidate matches between Cep-Her and the Kepler
objects of interest is as follows -- the objects are listed in order
of descending weights, $D$.
Objects designated as confirmed planets included
Kepler-1627,
Kepler-1643,
Kepler-1331,
Kepler-1062, and
Kepler-1933.  % Newly validated by Valizadegan+2022
Objects designated as candidate planets included
KOI-5264,
KOI-8007,
KOI-7572,
KOI-7375,
KOI-7368,
KOI-7638,
KOI-5632, and
KOI-7913.
Objects designated known false positive planet candidates included
KOI-6437, 
KOI-5988, 
KOI-7871,
KOI-7655,
KOI-5024,
KOI-61,
KOI-4336,
KOI-6812,
KOI-3399, and
KOI-6277.
Finally, Kepler-1902 (KOI-3090) has one confirmed planet
(KOI-3090.02), and one false positive (KOI-3090.01).
Of these objects, only Kepler-1627, Kepler-1643, KOI-7368, and
KOI-7913 met our requirements for potentially
both {\it i)} having real planets, and {\it ii)} being $\lesssim 10^8$
years old, based on the presence of rotational modulation at the
expected period and amplitude.  One object was ambiguous: Kepler-1933.
This system has a confirmed $\approx$$1.4\,R_\oplus$ planet, a stellar
rotation period of 6.5\ days, and an effective temperature of $\approx$$5750\,{\rm K}$.  This places it near the upper envelope of the
rotation period vs.\ color distribution for the Pleiades, making
it unlikely to be $\approx$40\,Myr old.  Nonetheless, we acquired a
reconnaissance HIRES spectrum, and it yielded ${\rm EW}_{\rm Li} = 93 \pm
5$\,m\AA.  Combined with the rotation period, this suggests an age
between 100 and 200\,Myr for Kepler-1933, but the system is
unlikely to be part of Cep-Her.

>>>>

% 208 words (250 max)
\begin{abstract}
  %The best way to measure a planet's age is to ignore the planet and
  %pay attention to the star.
  Stellar positions and velocities from Gaia are yielding a new window
  on open cluster dispersal.  
  Here we present an analysis of a group
  of stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
  hereafter the Cep-Her complex.  The group includes four Kepler
  Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in part of the Cep-Her complex
  called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.  Here
  we focus on the former three systems, which are in other regions of
  the association.  Based on kinematic evidence from Gaia, stellar
  rotation periods from TESS, and spectroscopy, these three objects
  are also $\approx$40 million years (Myr) old.  More specifically, we
  find that Kepler-1643 is $46^{+9}_{-7}$\,Myr old, based on its
  membership in a dense sub-cluster of the complex called RSG-5.
  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
  diffuse region that we call CH-2.  Based on the transit shapes and
  high resolution imaging, all three objects are most likely planets,
  with false positive probabilities of $6\times10^{-9}$,
  $4\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
  and KOI-7913 respectively.  These planets demonstrate
  that mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages
  of 40 million years.
\end{abstract}

@@

%
% KOI match numbers from calc_CepHer_group_numbers.py
%
Our fourth and final step was to cross-match the candidate Cep-Her
member list against all known Kepler Objects of Interest.  We used the
Cumulative KOI table from the NASA Exoplanet Archive from 27 March
2022, and also compared against the \texttt{q1\_q17\_dr25} table
\citep{thompson_planetary_2018}.  From the candidate members with
weights exceeding 0.02, this yielded 11 known false positives, 6
``confirmed'' planets, and 8 ``candidate'' planets.  Inspection of the
Kepler data validation summaries and Robovetter classifications for
these objects showed whether they were potentially consistent with
being {\it i)} planets, and {\it ii)} $\lesssim 10^8$ years old, based
on the presence of rotational modulation at the expected period and
amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}. 
Four
objects passed both cuts: Kepler-1627, Kepler-1643,
KOI-7368, and KOI-7913.

@@

\subsection{KOI-7368}
\paragraph{Spectra}
For KOI-7368, we acquired a spectrum on 2015 June 1 using the echelle
spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
Stellar Parameter Classification pipeline for TRES has been described
by \citet{2021tsc2.confE.124B}.  It is based on the synthetic template
library constructed by \citet{Buchhave2012}.  The resulting stellar
parameters ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
cluster-isochrone method within $1$-$\sigma$.  Auxiliary
spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
\pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm 0.50\,\mkms$,
and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
The Li 6708\AA\ EW measurement procedure yielded ${\rm EW}_{\rm Li} =
236^{+16}_{-14}$\,m\AA.

@@

The theoretical expectation is that mini-Neptunes with sizes of 2 to
3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
This expectation is tied to inferences about the initial distributions
of planetary core mass, core composition, and atmospheric mass
fraction \citep{owen_evaporation_2017}.  The Kelvin-Helmholtz cooling
timescale, which is tied to the entropy of the planetary interior
shortly after disk dispersal, also plays a significant role
\citep{owen_constraining_2020}.  As an example,
\citet{rogers_unveiling_2021} predicted that given a core mass
distribution peaked at $\approx$$4$\,$M_\oplus$, an ice-poor rock/iron
core composition, and a typical H/He mass fraction of $\approx$4\%,
there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
at times between 10 and 100\,Myr.  The models advanced by
\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} broadly agree;
their differences lie in the mechanism for producing the radius
valley, and in their assumed initial core mass distributions.

Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837, and TOI-1227
have sizes that are anomalously large relative to the predicted peak
in planet occurrence at 2 to 3\,$R_\oplus$.  However, their large
sizes can be accommodated by invoking any of {\it i)} larger core
masses, {\it ii)} more volatile-rich compositions, {\it iii)}  larger
initial atmospheric mass fractions, or {\it iv)} longer thermal
cooling times.  Secure mass measurements would help constrain this
parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
semi-amplitudes make measuring the Doppler orbits very difficult.
\citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
demonstrate that at least some planets at 40\,Myr have sizes that are
consistent with theoretical expectations for mini-Neptunes.  While
selection effects imposed by spot-induced photometric variability are
a likely explanation for why planets this small have not previously
been identified \citep[{e.g.},][]{zhou_2021_tois}, future work should
quantify this bias more carefully, in order to enable empirical
studies of how the planetary size distribution changes at early times.

@@

We have shown that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
to 50 million years old, and that each system is most likely
planetary.  The evidence for the planetary interpretation comes from
an application of \texttt{VESPA} to the Kepler data, alongside new
imaging from NIRC2.  The validity of the \texttt{VESPA} framework
rests on the premise that non-astrophysical false positives can be
rejected.  
This seems to be the case for all three objects, even though
Kepler-1643 and KOI-7368 
both show weak centroid offsets between the in
and out-of-transit phases.  For both cases, the weak observed shifts
are consistent with being caused by starspot-induced variability in
specific quarters spuriously moving the stellar center-of-light.
Independently, our imaging rules out companion stars with the
brightness and positions that would be needed to otherwise explain the
reported shifts.  All three objects are therefore validated planets.

@@

%\clearpage
\acknowledgements
\raggedbottom
% Cycle 3 and 4
L.G.B{.} is supported by the Heising-Simons Foundation 51
Pegasi~b Fellowship and the NASA TESS GI Program (80NSSC21K0335
and 80NSSC22K0298).
R.K{.} is supported by the Heising-Simons Foundation.
J.L.C{.} is supported by NSF AST-2009840 and the NASA TESS GI Program
(80NSSC22K0299).
D.H. is supported by the Alfred P. Sloan Foundation and NASA (80NSSC19K0597).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%%
%We thank the TESS architects
%%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
%and the many TESS team members for their
%efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
%Finally, this research has made use of the Keck Observatory Archive (KOA),
%which is operated by the W. M. Keck Observatory and the NASA Exoplanet
%Science Institute (NExScI), under contract with the National
%Aeronautics and Space Administration.
%We also thank the Keck Observatory staff for their support of
%HIRES and remote observing.  We recognize the importance that the
%summit of Maunakea has within the indigenous Hawaiian community, and
%are deeply grateful to have the opportunity to conduct observations
%from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\paragraph{Table~2} contains 338 candidate Cep-Her members with weights
$D>0.02$ observed by Kepler.  The final catalog of candidate
Cep-Her members will be provided by R.~Kerr et al.\ in prep; Table~2 is
from a preliminary version of that work.  Note that more
restrictive weight cuts should be imposed if one wishes to remove
the majority of field star interlopers.
Table~2 was created by cross-matching candidate Cep-Her members
(selected using Gaia EDR3; Section~\ref{subsec:members}) against a
Kepler to Gaia DR2 cross-match (the \texttt{gaia-kepler.fun}
crossmatch database created by Megan Bedell).  The
\texttt{kic\_dr2\_ang\_dist} column is from the latter table.  The
EDR3 to DR2 match was performed using the
\texttt{gaiaedr3.dr2\_neighbourhood} table, and the closest proper
motion and epoch-corrected angular distance neighbor was taken as the
single best match.  The \texttt{edr3\_dr2\_mag\_diff} column gives
some indication of the reliability of this EDR3 to DR2 conversion, as
there are a few cases between Gaia DR2 and EDR3 where partially
resolved binaries became fully resolved.

\paragraph{Candidate matches between Cep-Her and the Kepler Objects of
Interest:}
The full list of candidate matches between Cep-Her and the Kepler
Objects of Interest is as follows -- the objects are listed in order
of descending weights, $D$.
Objects designated as confirmed planets included
Kepler-1627,
Kepler-1643,
Kepler-1331,
Kepler-1062, and
Kepler-1933.  % Newly validated by Valizadegan+2022
Objects designated as candidate planets included
KOI-5264,
KOI-8007,
KOI-7572,
KOI-7375,
KOI-7368,
KOI-7638,
KOI-5632, and
KOI-7913.
Objects designated known false positive planet candidates included
KOI-6437, 
KOI-5988, 
KOI-7871,
KOI-7655,
KOI-5024,
KOI-61,
KOI-4336,
KOI-6812,
KOI-3399, and
KOI-6277.
Finally, Kepler-1902 (KOI-3090) has one confirmed planet
(KOI-3090.02), and one false positive (KOI-3090.01).
Of these objects, only Kepler-1627, Kepler-1643, KOI-7368, and
KOI-7913 met our requirements for potentially
both {\it i)} having real planets, and {\it ii)} being $\lesssim 10^8$
years old, based on the presence of rotational modulation at the
expected period and amplitude.  
Of the
14 confirmed and candidate planets,
6 failed first filter, and 7 independently failed the second.
One object was ambiguous: Kepler-1933.
This system has a confirmed $\approx$$1.4\,R_\oplus$ planet, a stellar
rotation period of 6.5\ days, and an effective temperature of $\approx$$5750\,{\rm K}$.  This places it near the upper envelope of the
rotation period vs.\ color distribution for the Pleiades, making
it unlikely to be $\approx$40\,Myr old.  Nonetheless, we acquired a
reconnaissance HIRES spectrum, and it yielded ${\rm EW}_{\rm Li} = 93 \pm
5$\,m\AA.  Combined with the rotation period, this suggests an age
for Kepler-1933 between 100 and 300\,Myr.  Based on these indicators, the system is
unlikely to be part of Cep-Her, but could merit further study.
